### 2020-08-24 ART_Piechowicz/Piechowicz-PU.tex
A. Figarski correctons added. Sending to the authors

\indent Like in many other issues strong background for discussion of the notions of knowledge and belief from the logical point of view was provided by Aristotle in his works such as \textit{De Sophisiticis Elenchis} and both \textit{Analytics}. His analysis concerned the concepts of possibility, necessity, impossibility and contingency  became the important base for medieval thinkers such as Buridan, Pseudo Scotus, Ockham, and Ralph Strode who have extended the Aristotelian concepts to epistemic themes and problems \parencites[see][]{boh_epistemic_1993}{knuuttila_modalities_1993}.
%(see Boh, 1993; Knuuttilla, 1993).
During this period, the Pseudo-Scot and William of Ockham supplemented Aristotle’s study of mental acts of cognition and volition
\parencite[see][p.130]{boh_epistemic_1993}.
%(see Boh 1993, p. 130).
Topics of epistemic logic discussed by medieval thinkers have a similar set of foundational assumptions with modern discussions like connection between knowledge and veracity and an observation that an epistemic agent cannot coherently assert ``p but I do not believe (know) p'', which was explicitly stated  by G. E. Moore in the 20\textsuperscript{th} century.

@@

The system of Jerzy \L{}o\'{s} was truly pioneering and very interesting because it fixed quite strong meaning of the functor of assertion. First, assertion fixes non-contradictorily, that is, if an epistemic subject asserts a sentence then he can’t asserts its negation. This requirement is acceptable in the systems of the normal modal logic but many thinkers discuss its legitimacy applied to the real convictions. Moreover, this axiom asserts the completeness of system of beliefs which is more controversial. Despite of this---as it was mentioned in first part of this article---the first axiom of \L{}o\'{s}’s system---like its equivalents in other logics---designates the meaning of the functor of assertion and is not a description of its everyday use.\footnote{Those problems are discussed in
\parencites[][p.80-82, 93-94]{marciszewski_podstawy_1972}{poczobut_sprzecznosci_1999}.}
%(Marciszewski, 1972, p. 80-82, 93-94; Poczobut, 1997).}
%tu

@@

A more liberal system may be generated by the weakening by the first of the axioms, that is\footnote{This question (and many others) are broadly discuted in
\parencites[][p.254-264]{lechniak_przekonania_2011}[][p.75-88]{marciszewski_podstawy_1972}.}:
%(Lechniak, 2011, p. 254-264; Marciszewski, 1972, p. 75-88).}:
$$\mathbf{L}_{x}p\rightarrow \neg \mathbf{L}_{x}\neg p.$$
What are the consequences of this modification? In system of \L{}o\'{s} the assertion of some state implies that its negation is not asserted and, equivalently, if negation is asserted then this state is not. The second implication is more problematic than the first because of the formulas:
$$\neg \mathbf{L}_{x}\neg p\rightarrow  \mathbf{L}_{x} p$$
$$\mathbf{L}_{x}\neg p\vee \mathbf{L}_{x}p,$$
$$\mathbf{L}_{x}(\mathbf{L}_{x}\neg p\vee \mathbf{L}_{x}p).$$
The last of these formulas generated only by definability of functors and last axiom is a declaration of the omniscience of a subject. There is no place to this problematic thesis in the modified system.

@@

\indent Although the doxastic logics are well-founded systems today, the complete description of the cognitive capacity of human beings (or AI) should include language ability because beliefs are epistemically inaccessible without communication. This statement leads to the non-trivial problem of the formal description of how the coding of beliefs in utterances is accomplished. One way of solving this problem requires the formalization of some informal theory of communication and the best candidate for this purpose was the theory formulated by P.H. Grice
\parencite*[p.22-40]{grice_studies_1989}.
%(1989, p. 22--40).
The theory of the conversational implicatures was proposed by him to explain what people have said, that is, what is the meaning of the sentences uttered by them in a concrete context.

>>>>

\indent Like in many other issues strong background for discussion of the notions of knowledge and belief from the logical point of view was provided by Aristotle in his works such as \textit{De Sophisiticis Elenchis} and both \textit{Analytics}. His analysis concerned the concepts of possibility, necessity, impossibility and contingency  became the important base for medieval thinkers such as Buridan, Pseudo Scotus, Ockham, and Ralph Strode who have extended the Aristotelian concepts to epistemic themes and problems \parencites[see][]{boh_epistemic_1993}{knuuttila_modalities_1993}.
%(see Boh, 1993; Knuuttilla, 1993).
During this period, the Pseudo-Scot and William of Ockham supplemented Aristotle’s study of mental acts of cognition and volition
\parencite[see][p.130]{boh_epistemic_1993}.
%(see Boh 1993, p. 130).
Topics of epistemic logic discussed by medieval thinkers have a similar set of foundational assumptions with modern discussions like connection between knowledge and veracity and an observation that an epistemic agent cannot coherently assert ``p but I do not believe (know) p'', which was explicitly stated  by G.E. Moore in the 20\textsuperscript{th} century.

@@

The system of Jerzy \L{}o\'{s} was truly pioneering and very interesting because it fixed quite strong meaning of the functor of assertion. First, assertion fixes non-contradictorily, that is, if an epistemic subject asserts a sentence then he can’t asserts its negation. This requirement is acceptable in the systems of the normal modal logic but many thinkers discuss its legitimacy applied to the real convictions. Moreover, this axiom asserts the completeness of system of beliefs which is more controversial. Despite of this---as it was mentioned in first part of this article---the first axiom of \L{}o\'{s}’s system---like its equivalents in other logics---designates the meaning of the functor of assertion and is not a description of its everyday use.\footnote{Those problems are discussed in
\parencites[][pp.80-82, 93-94]{marciszewski_podstawy_1972}{poczobut_sprzecznosci_1999}.}
%(Marciszewski, 1972, p. 80-82, 93-94; Poczobut, 1997).}
%tu

@@

A more liberal system may be generated by the weakening by the first of the axioms, that is\footnote{This question (and many others) are broadly discuted in
\parencites[][pp.254-264]{lechniak_przekonania_2011}[][pp.75-88]{marciszewski_podstawy_1972}.}:
%(Lechniak, 2011, p. 254-264; Marciszewski, 1972, p. 75-88).}:
$$\mathbf{L}_{x}p\rightarrow \neg \mathbf{L}_{x}\neg p.$$
What are the consequences of this modification? In system of \L{}o\'{s} the assertion of some state implies that its negation is not asserted and, equivalently, if negation is asserted then this state is not. The second implication is more problematic than the first because of the formulas:
$$\neg \mathbf{L}_{x}\neg p\rightarrow  \mathbf{L}_{x} p$$
$$\mathbf{L}_{x}\neg p\vee \mathbf{L}_{x}p,$$
$$\mathbf{L}_{x}(\mathbf{L}_{x}\neg p\vee \mathbf{L}_{x}p).$$
The last of these formulas generated only by definability of functors and last axiom is a declaration of the omniscience of a subject. There is no place to this problematic thesis in the modified system.

@@

\indent Although the doxastic logics are well-founded systems today, the complete description of the cognitive capacity of human beings (or AI) should include language ability because beliefs are epistemically inaccessible without communication. This statement leads to the non-trivial problem of the formal description of how the coding of beliefs in utterances is accomplished. One way of solving this problem requires the formalization of some informal theory of communication and the best candidate for this purpose was the theory formulated by P.H. Grice
\parencite*[pp.22-40]{grice_studies_1989}.
%(1989, p. 22--40).
The theory of the conversational implicatures was proposed by him to explain what people have said, that is, what is the meaning of the sentences uttered by them in a concrete context.
